abxD01 Paper

### 2015-04-10 Schubert_abxD01_mBio_2015.Rmd
Format document

**Antibiotics differentially alter the structure of the microbiota and their colonization resistance to** ***C. difficile.*** We selected a panel of seven antibiotics from six classes with the goal of differentially altering the microbiota and assessing their resistance to *C. difficile* colonization (**Table 1**). On the day after mice were challenged with *C. difficile* we enumerated the density of *C. difficile* in the animals' feces. We observed reproducibly high levels of *C. difficile* colonization in mice treated with cefoperazone, metronidazole, and streptomycin (**Figures 1 and S1**). We observed variable levels of *C. difficile* colonization in mice treated with ampicillin. Only one of six mice receiving vancomycin was colonized with *C. difficile* and none of the mice that received ciprofloxacin were colonized. We sequenced the 16S rRNA genes from the fecal communities of treated and untreated mice prior to *C. difficile* challenge to identify populations within the microbiota that conferred colonization resistance. All of the antibiotic treatments, except for the ciprofloxacin-treated mice (AMOVA, , P=`r round(max(control_sig[control_comparison=="cipro-control"]), 2)`), resulted in distinct and reproducible changes to the structure of the microbiota relative to the untreated animals (AMOVA, P<`r ifelse(max(control_sig[control_comparison!="cipro-control"]) < 1e-3, "0.001", "NA")`). Comparisons of the microbiota between antibiotic classes indicated that their structures were significantly different from each other (AMOVA, P<`r round(max(other_sig[other_comparison!="amp-cef"]), 2)`). The community structures of mice receiving beta-lactams (i.e. cefoperazone and ampicillin) were not significantly different from each other (AMOVA, P=`r round(max(other_sig[other_comparison=="amp-cef"]), 2)`). These results indicate that perturbing the gut microbiota with antibiotics resulted in non-overlapping community structures that yielded significant variation in susceptibility to colonization when challenged with *C. difficile*.

@@

There is clear need for more efficient therapies for treatment of *C. difficile* infections in humans aimed at restoration of the microbiota. Current first line treatments of CDI include regimens of either metronidazole or vancomycin, which further perturb the microbiota. As such, relapse rates for CDI are typically around 25-30% {Wilcox, 1998 \#599}. Interestingly, we observed that treatment with either antibiotic induced susceptibility to *C. difficile* in mice. This result has implications for understanding the causes of recurrent infections. Previous efforts to restore the microbiota and reestablish colonization resistance also support our findings. For instance, association of germ-free mice with a *Lachnospiraceae* isolate only reduced the level of *C. difficile* colonization by 10 to 100-fold {Reeves, 2012 \#126}. Using conventional mice, mixtures of bacteria rather than individual populations have been shown to restore colonization resistance and mediate clearance of *C. difficile* {Lawley, 2012 \#36; Buffie, 2015 \#981}. Fecal transplants, which represent a diverse collection of bacterial populations, have been highly effective in treating humans with recurrent *C. difficile* {Kassam, 2013 \#1003; Seekatz, 2014 \#1051; Weingarden, 2015 \#1054}. By generating a diverse collection of communities that were challenged with **C. difficile**, we have identified a subset of populations using random forest modeling that could be used as a probiotic cocktail to provide colonization resistance. These would include members of the *Porphyromonadaceae*, *Lachnospiraceae*, *Lactobacillus*, and *Alistipes*. Several of these populatins have been examined for their potential as a probiotic for preventing *C. difficile* infection. A 6-species cocktail, including isolates of *Porphyromonadaceae*, *Lachnospiraceae*, *Lactobacillus*, *Coriobacteriaceae*, *Staphylococcus*, and *Enterococcus*, successfully resolved CDI in mice {Lawley, 2012 #36}. In humans, *Lactobacillus*-based probiotics have been co-administered with antibiotics to deter the onset of antibiotic-associated diarrhea (AAD) and *C. difficile* infection {Pattani, 2013 #1004}. A more diverse probiotic contained 33 bacterial species included *Porphyromonadaceae*, *Lachnospiraceae*, *Ruminococcaeae*, *Eubacteriaceae*, and *Lactobacillus* isolates, successfully restored colonization resistance in recurrent *C. difficile* infection and eliminated diarrhea up to 6 months post treatment {Petrof, 2013 #2}. Given this evidence, we feel confident that an effective probiotic community could be designed based on our findings to recover colonization resistance against **C. difficile**.

@@

| Antibiotic | Administration | Class | Mechanism | Target |
|:---------:|:--------------:|:-----:|:---------:|:---------:|
Ciprofloxacin | Oral gavage, one time | Fluoroquinolone | Inhibits DNA gyrase | Gram +/-
Clindamycin | IP injection, one time | Lincosamide | Inhibits protein synthesis | Anaerobes
Vancomycin | *Ad libitum* in drinking water, five days | Glycopeptide | Inhibits peptidoglycan synthesis | Gram +
Streptomycin | *Ad libitum* in drinking water, five days | Aminoglycoside | Inhibits protein synthesis | Gram +/-
Cefoperazone | *Ad libitum* in drinking water, five days | β-lactam: Cephalosporin | Inhibits peptidoglycan synthesis | Gram +/-
Ampicillin | *Ad libitum* in drinking water, five days | β-lactam: Penicillin | Inhibits peptidoglycan synthesis | Gram +/-
Metronidazole | *Ad libitum* in drinking water, five days | Nitromidazole | Destabilizes bacterial DNA | Anaerobes

@@

**Figure 6. Relationship between OTU relative abundance and** ***C. difficile*** **colonization levels indicates non-linearity and context-dependency.** The `r n_part_features` OTUs that resulted in the greatest change in percent mean squared error when removed from the random forest regression model are shown in each pane and together explain `r round(rsq_part_features, 1)`% of the variation in the data. The Spearman correlation value between that OTUs abundance and **C. difficile** levels are shown for each pane. The color and symbols represent the same antibiotic dose and recovery period as in Figure 5.

>>>>

**Antibiotics differentially alter the structure of the microbiota and their colonization resistance to** ***C. difficile.*** We selected a panel of seven antibiotics from six classes with the goal of differentially altering the microbiota and assessing their resistance to *C. difficile* colonization (**Table 1**). On the day after mice were challenged with *C. difficile* we enumerated the density of *C. difficile* in the animals' feces. We observed reproducibly high levels of *C. difficile* colonization in mice treated with cefoperazone, metronidazole, and streptomycin (**Figures 1 and S1**). We observed variable levels of *C. difficile* colonization in mice treated with ampicillin. Only one of six mice receiving vancomycin was colonized with *C. difficile* and none of the mice that received ciprofloxacin were colonized. We sequenced the 16S rRNA genes from the fecal communities of treated and untreated mice prior to *C. difficile* challenge to identify populations within the microbiota that conferred colonization resistance. All of the antibiotic treatments, except for the ciprofloxacin-treated mice (AMOVA, P=`r round(max(control_sig[control_comparison=="cipro-control"]), 2)`), resulted in distinct and reproducible changes to the structure of the microbiota relative to the untreated animals (AMOVA, P<`r ifelse(max(control_sig[control_comparison!="cipro-control"]) < 1e-3, "0.001", "NA")`). Comparisons of the microbiota between antibiotic classes indicated that their structures were significantly different from each other (AMOVA, P<`r round(max(other_sig[other_comparison!="amp-cef"]), 2)`). The community structures of mice receiving beta-lactams (i.e. cefoperazone and ampicillin) were not significantly different from each other (AMOVA, P=`r round(max(other_sig[other_comparison=="amp-cef"]), 2)`). These results indicate that perturbing the gut microbiota with antibiotics resulted in non-overlapping community structures that yielded significant variation in susceptibility to colonization when challenged with *C. difficile*.

@@

There is clear need for more efficient therapies for treatment of *C. difficile* infections in humans aimed at restoration of the microbiota. Current first line treatments of CDI include regimens of either metronidazole or vancomycin, which further perturb the microbiota. As such, relapse rates for CDI are typically around 25-30% {Wilcox, 1998 \#599}. Interestingly, we observed that treatment with either antibiotic induced susceptibility to *C. difficile* in mice. This result has implications for understanding the causes of recurrent infections. Previous efforts to restore the microbiota and reestablish colonization resistance also support our findings. For instance, association of germ-free mice with a *Lachnospiraceae* isolate only reduced the level of *C. difficile* colonization by 10 to 100-fold {Reeves, 2012 \#126}. Using conventional mice, mixtures of bacteria rather than individual populations have been shown to restore colonization resistance and mediate clearance of *C. difficile* {Lawley, 2012 \#36; Buffie, 2015 \#981}. Fecal transplants, which represent a diverse collection of bacterial populations, have been highly effective in treating humans with recurrent *C. difficile* {Kassam, 2013 \#1003; Seekatz, 2014 \#1051; Weingarden, 2015 \#1054}. By generating a diverse collection of communities that were challenged with *C. difficile*, we have identified a subset of populations using random forest modeling that could be used as a probiotic cocktail to provide colonization resistance. These would include members of the *Porphyromonadaceae*, *Lachnospiraceae*, *Lactobacillus*, and *Alistipes*. Several of these populatins have been examined for their potential as a probiotic for preventing *C. difficile* infection. A 6-species cocktail, including isolates of *Porphyromonadaceae*, *Lachnospiraceae*, *Lactobacillus*, *Coriobacteriaceae*, *Staphylococcus*, and *Enterococcus*, successfully resolved CDI in mice {Lawley, 2012 #36}. In humans, *Lactobacillus*-based probiotics have been co-administered with antibiotics to deter the onset of antibiotic-associated diarrhea (AAD) and *C. difficile* infection {Pattani, 2013 #1004}. A more diverse probiotic contained 33 bacterial species included *Porphyromonadaceae*, *Lachnospiraceae*, *Ruminococcaeae*, *Eubacteriaceae*, and *Lactobacillus* isolates, successfully restored colonization resistance in recurrent *C. difficile* infection and eliminated diarrhea up to 6 months post treatment {Petrof, 2013 #2}. Given this evidence, we feel confident that an effective probiotic community could be designed based on our findings to recover colonization resistance against *C. difficile*.

@@

| Antibiotic | Administration | Class | Mechanism | Target |
|:---------:|:--------------:|:-----:|:---------:|:---------:|
| Ciprofloxacin | Oral gavage, one time | Fluoroquinolone | Inhibits DNA gyrase | Gram +/- |
| Clindamycin | IP injection, one time | Lincosamide | Inhibits protein synthesis | Anaerobes |
| Vancomycin | *Ad libitum* in drinking water, five days | Glycopeptide | Inhibits peptidoglycan synthesis | Gram + |
| Streptomycin | *Ad libitum* in drinking water, five days | Aminoglycoside | Inhibits protein synthesis | Gram +/- |
| Cefoperazone | *Ad libitum* in drinking water, five days | β-lactam: Cephalosporin | Inhibits peptidoglycan synthesis | Gram +/- |
| Ampicillin | *Ad libitum* in drinking water, five days | β-lactam: Penicillin | Inhibits peptidoglycan synthesis | Gram +/- |
| Metronidazole | *Ad libitum* in drinking water, five days | Nitromidazole | Destabilizes bacterial DNA | Anaerobes |

@@

**Figure 6. Relationship between OTU relative abundance and** ***C. difficile*** **colonization levels indicates non-linearity and context-dependency.** The `r n_part_features` OTUs that resulted in the greatest change in percent mean squared error when removed from the random forest regression model are shown in each pane and together explain `r round(rsq_part_features, 1)`% of the variation in the data. The Spearman correlation value between that OTUs abundance and *C. difficile* levels are shown for each pane. The color and symbols represent the same antibiotic dose and recovery period as in Figure 5.
